si: code and prose exploring Spatial Interaction models

### 2022-04-15 README.qmd
Add readme with minimal example and description for #1
Former-commit-id: f2a35a8d2dac6764957ca2bd1b7c32a44e3ac588

---
# format: html
format: gfm
number-sections: true
title: "si: code and prose exploring Spatial Interaction models"
bibliography: vignettes/references.bib
execute:
  cache: false
  fig.path: "man/figures/README-"
---

@@

The goal of {si} is to demonstrate Spatial Interaction models from first principles (see the `sims` vignette for a more detailed description of SIMs and implementation in R without using this package for reference).
The package also contains reproducible code and documentation demonstrating how they work.

@@

-   Unconstrained

-   Production-constrained

-   Attraction-constrained

-   Doubly-constrained

@@

# Implementations in other languages

@@

# Resources

>>>>

---
# format: html
format: gfm
title: "si: code and prose exploring Spatial Interaction models"
bibliography: vignettes/references.bib
execute:
  cache: false
  fig.path: "man/figures/README-"
---

@@

The goal of {si} is to demonstrate Spatial Interaction models from first principles (see the [`sims` vignette](https://robinlovelace.github.io/si/articles/sims.html) for a more detailed description of SIMs and implementation in R without using this package for reference).
The package also contains reproducible code and documentation demonstrating how they work.

@@

-   Unconstrained
-   Production-constrained
-   Attraction-constrained
-   Doubly-constrained

@@

<!-- # Implementations in other languages -->
## si basics

Install the package as follows:

```r
install.packages("remotes") # if not already installed
```

```{r installation}
remotes::install_github("robinlovelace/si")
```

Run a basic SIM as follows:

```{r distance plot}
library(si)
od = si_to_od(origins = zones, destinations = zones, max_euclidean_distance = 5000)
gravity_model = function(od, beta) {
  od[["origin_all"]] * od[["destination_all"]] *
    exp(-beta * od[["distance_euclidean"]] / 1000)
} 
od_res = si_model(od, fun = gravity_model, beta = 0.3, var_p = origin_all)
plot(od_res$distance_euclidean, od_res$res)
```

What just happened?
As the example above shows, the package allows/encourages you to define and use your own functions to estimate the amount of interaction/movement between places.
The resulting estimates of interaction, returned in the column `res` and plotted with distance in the graphic above, resulted from our choice of spatial interaction model inputs, allowing a wide range of alternative approaches to be implemented.
This flexibility is a key aspect of the package, enabling small and easily modified functions to be implemented and tested.

The output `si_model()` is a geographic object which can be plotted as a map:

```{r map}
plot(od_res["res"], logz = TRUE)
```

Perhaps the most important function in the package is therefore `si_to_od()`, which transforms geographic entities (typically zones but points and other geographic features could be used) into a data frame representing the full combination of origin-destination pairs that are less than `max_euclidean_distance` meters apart.
A common saying in data science is that 80% of the effort goes into the pre-processing stage.
This is equally true for spatial interaction modelling as it is for other types of data intensive analysis/modelling work.
So what does this function return?

```{r}
names(od)
```

As shown in the output above, the function allows you to use any variable in the origin or destination data in the function, with column names appended with `origin` and `destination`.

Note: support for OD datasets in which the origins and destinations are different objects is work in progress (see [#3](https://github.com/Robinlovelace/si/issues/3)).

## Feedback

@@

## Alternative approaches

The {si} package is nascent. Anyone looking for more mature approaches could check the following:
